---
title: "ggblanket: some basics"
author: "David Hodge"
date: "2023-05-03"
slug: "ggblanket-why"
categories: R
tags: R
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

ggblanket seeks to simplify ggplot2 visualisation.

The below article outlines some basics of how the package works.

```{r}
library(tidyverse)
library(ggblanket)
library(patchwork)
library(palmerpenguins)

penguins <- penguins |>
  mutate(sex = stringr::str_to_sentence(sex)) |>
  tidyr::drop_na(sex)
```

1.  Over thirty gg_* wrapper functions to plot a single geom

```{r}
penguins |> 
  ggplot() + 
  geom_point(aes(x = flipper_length_mm, y = body_mass_g))

penguins |>
  gg_point(
    x = flipper_length_mm,
    y = body_mass_g)
```

2.  A single col argument to colour and fill by a variable

```{r}
penguins |>
  ggplot() +
  geom_density(aes(x = body_mass_g, fill = species))

penguins |>
  gg_density(
    x = body_mass_g, 
    col = species)
```

3.  unifed approach to changing colours via a pal argument

```{r}
penguins |>
  ggplot() +
  geom_histogram(
    aes(x = body_mass_g),
    fill = "#800000")

penguins |>
  gg_histogram(
    x = body_mass_g, 
    pal = "#800000")
```

```{r}
penguins |>
  ggplot() +
  geom_jitter(aes(x = species, y = body_mass_g, col = sex)) +
  scale_color_manual(values = c("#fa8072", "#468499"))

penguins |>
  gg_jitter(
    x = species, 
    y = body_mass_g, 
    col = sex, 
    pal = c("#fa8072", "#468499"))
```

4.  A facet argument to facet by a variable

```{r}
penguins |>
  ggplot() +
  geom_violin(aes(x = sex, y = body_mass_g)) +
  facet_wrap(~species) 

penguins |>
  gg_violin(
    x = sex,
    y = body_mass_g,
    facet = species)
```

5. An additional facet2 argument to for grid faceting

```{r, fig.height=5}
penguins |>
  ggplot() +
  geom_histogram(aes(x = flipper_length_mm)) +
  facet_grid(rows = vars(species), cols = vars(sex))

penguins |>
  gg_histogram(
    x = flipper_length_mm,
    facet = sex,
    facet2 = species)
```

6.  prefixed arguments easy to find using autotab

```{r}
penguins |>
  ggplot() +
  geom_boxplot(aes(x = sex, y = body_mass_g, fill = sex), 
               show.legend = F) +
  facet_wrap(~species) +
  expand_limits(y = 0) +
  scale_x_discrete(labels = \(x) stringr::str_sub(x, 1, 1)) +
  scale_y_continuous(breaks = scales::breaks_width(500))

penguins |>
  gg_boxplot(
    x = sex,
    y = body_mass_g,
    col = sex,
    facet = species, 
    x_labels = \(x) stringr::str_sub(x, 1, 1),
    y_include = 0, 
    y_breaks = scales::breaks_width(500))

```

7. Default titles converted to sentence case

```{r}
penguins |> 
  group_by(species, sex) |> 
  summarise(body_mass_g = mean(body_mass_g, na.rm = TRUE)) |> 
  ggplot() +
  geom_col(aes(x = body_mass_g, y = species, fill = sex),
           position = "dodge",
           width = 0.75) +
  labs(y = "Penguin species")  

penguins |> 
  group_by(species, sex) |> 
  summarise(body_mass_g = mean(body_mass_g, na.rm = TRUE)) |> 
  gg_col(
    x = body_mass_g, 
    y = species, 
    col = sex, 
    position = "dodge",
    width = 0.75,
    y_title = "Penguin species")  
```

8.  default numeric scales working in pretty ways

Where the orientation is normal (i.e. vertical): 

* Default x scale limits and expanding are as per ggplot2 defaults
* Default numeric y scale limits start and end on a break with zero expanding

It does the opposite where the orientation is horizontal.

Default gridllines work with this, and the effect is that the plot looks very nice and symmetrical - to my eye anyway! 




For further information, see the [ggblanket website](https://davidhodge931.github.io/ggblanket/).
